An Efficient Parameter Space Search as an Alternative to Markov Chain Monte Carlo

### 2014-07-21 aps4p5/drafts/draft_140719.tex
more writing

Gaussian processes use the sampled data $\{\vec{\theta}^{(i)},\chi^{2,(i)}\}$, 
where $i$ indexes over all of the sampled points, to predict
$\mu=\chi^{2,(q)}$ at the candidate point 
$\vec{\theta}^{(q)}$ by assuming that the function
$\vec{\theta}\rightarrow\chi^2$ represents a sample drawn from a random process 
distributed
across parameter space.  At each point in parameter space, 
$\chi^2$ is assumed to be
distributed according to a normal distribution with mean 
$\bar{\chi}^2$
and variance dictated by the covariance function 
$C_{ij}(\vec{\theta}^{(i)},\vec{\theta}^{(j)})$.
If we have $N_g$ measurements of $\chi^2$, then we assume
they are distributed according to
\begin{equation}
\label{eqn:likelihood}
P(\vec{\chi}^2)=
\frac{\exp\bigg[-\frac{1}{2}K^{-1}\sum_{i,j}^{N_g}
(\chi^{2,(i)}-\bar{\chi}^2)C^{-1}_{ij}
(\chi^{2,(j)}-\bar{\chi}^2)\bigg]}{\sqrt{2\pi}^{N_g}\text{det}|KC|^{N_g/2}}
\end{equation}
$C_{ij}$ is a function assumed by the user encoding how variations in
$\chi^2$ at one point in parameter space affect variations in 
$\chi^2$ at other
points in parameter space.  $K$ is a parameter controlling the
normalization of $C_{ij}$.
The diagonal elements $C_{ii}=1+\sigma^2_{ii}$ where 
$\sigma^2_{ii}$ is the intrinsic variance in the value
of $\chi^2$ at a single point $\vec{\theta}^{(i)}$.  
Rasmussen and Williams (2006) treat 
the special case
$\bar{\chi}^2=0$ and find (see their equations 2.19, 2.25 and 2.26)
\begin{eqnarray}
\mu&=&\sum_{i,j} C_{qi}C^{-1}_{ij}\chi^{2,(j)}\nonumber\\
\sigma^2&=&C_{qq}-\sum_{i,j}C_{qi}C^{-1}_{ij}C_{jq}\label{sig}
\end{eqnarray}
where the sums over $i$ and $j$ are sums over the sampled points 
in parameter space.  $C_{iq}$ relates the sampled point $i$ to the candidate
point $q$.
We do not wish to assume that the mean value of $\chi^2$ is zero everywhere.
Therefore, we modify the equation for $\mu$ to give
\begin{equation}
\label{mu}
\mu=\bar{\chi}^2+\sum_{i,j} C_{qi}C^{-1}_{ij}
(\chi^{2,(j)}-\bar{\chi}^2)
\end{equation}
where $\bar{\chi}^2$ is the algebraic mean of the 
sampled $\chi^{2,(i)}$.
Note the similarity to a multi-dimensional Taylor series expansion with the
covariance matrix playing the role of the derivatives.
Equation (\ref{sig}) differs from equation (6) in \cite{bryan} because
they used the semivariance
$$\gamma_{ij}=\text{var}[\chi^2(\vec{\theta}^{(i)})
-\chi^2(\vec{\theta}^{(j)})]$$
in place of the covariance $C_{ij}$.  In practice, the two assumptions result in
equivalently valid $\mu$ and $\sigma$.

@@

@@

The parameter space we explore is the six-dimensional space of the spatially
flat concordance $\Lambda$CDM cosmology:
\begin{itemize}
\item$\Omega_bh^2$ -- the density of baryons in the Universe
\\
\item$\Omega_\text{CDM}h^2$ -- the density of cold dark matter in the Universe
\\
\item$\tau$ -- the optical depth to the surface of last scattering.  $\tau$ cannot be
constrained without polarizationd data, which we do not consider here.  However, it functions
as a nuisance parameter to be marginalized over.  The fact that APS can handle it as well as
MCMC is an important test of APS' feasibility.
\\
\item$h$ -- the Hubble parameter in units of
$100~\text{km}~\text{s}^{-1}~\text{Mpc}^{-1}$
\\
\item$n_s$ -- the spectral index controlling the distribution of scalar density
perturbations in the Universe
\\
\item$\ln[10^{10}A_s]$ -- the amplitude of the initial scalar density
perturbations in the Universe
\end{itemize}
These are the parameters taken by the publicly available MCMC code CosmoMC,
which we use to run our baseline MCMC chains \cite{cosmomc}.  The
$\vec{\theta}\rightarrow\chi^2$ function in this case involves converting these
paramters into the power spectrum of anisotropy $C_\ell$s on the sky and
comparing those $C_\ell$s to the actual measurements registered by the WMAP
satellite.  Using the (very fast) Boltzmann code CAMB to calculate the $C_\ell$s
\cite{camb} takes 1.3 seconds on a 2.5 GHz processor.  
Evaluating the $\chi^2$ using the likelihood code provided by the WMAP team
\cite{wmap7likelihood} takes an additional 2 seconds.  Hence the need
for an algorithm like APS or MCMC to accelerate the search.
We run APS with the following input parameters:
\begin{eqnarray}
\Delta\chi^2&=&12.59 \nonumber\\
&&\text{(according to the likelihood ratio test,}\nonumber\\
&&\text{the
95\% confidence limit for a 6-dimensional}\nonumber\\
&&\text{ parameter space
ought to be }\chi^2_\text{min}+12.59\text{)}\nonumber\\
R_g&=&0.1\nonumber\\
N_w&=&100\nonumber\\
N_c&=&250\nonumber\\
N_g&=&15\nonumber
\end{eqnarray}
We will consider both equations (\ref{eqn:covraw}) and (\ref{eqn:cov_nn})
as covariograms, though we find that the choice has little effect on our
results.

@@

The second question we consider is whether or not our assumption that
$\vec{\theta}\rightarrow\chi^2$ can be modeled by a Gaussian Process is 
valid.  In Figure
\ref{fig:goodness} we plot the average fractional error in $\mu$ as a proxy for $\chi^2$ as a
function of the number of steps sampled (note: the vertical axis actually shows the average
fractional error over the preceding 2,000 steps so that the early, imprecise models do not
contaminate the later, precise models).  As APS samples more points, its Gaussian Process
model becomes more accurate and the fractional error converges to a few percent.  Thus our
use of Gaussian Processes to model $\vec{\theta}\rightarrow\chi^2$ is also justified.

@@

\begin{figure}
\includegraphics[scale=0.3]{goodness_of_fit.eps}
\caption{
The average value of $|\mu-\chi^2|/\chi^2$ as a function of the number of points
sampled by APS in the WMAP 7 problem.  
This is shown as a check on our assumption that
$\vec{\theta}\rightarrow\chi^2$ can be approximated as a Gaussian Process.
As the number of sample points increases, the average error yielded by either
covariogram quickly converges to a few percent in $\chi^2$.
}
\label{fig:goodness}
\end{figure}

We now consider the ultimate results, namely the 95\% confidence limits on
our 6-dimensional cosmological parameter space yielded by APS.
To provide a baseline against which to test APS, we run 4 independent MCMC
chains using CosmoMC.  Like the MCMC code used in Section \ref{sec:toy},
CosmoMC periodically adjusts its proposal density by learning the covariance
matrix of the points it has already sampled.  At each step, it proposes
a new point in parameter space by randomly selecting an eigen vector
of that learned covariance matrix and stepping along it.
We integrate the posterior by taking the output chains, discarding the first
50\% of steps as a burn in period, and thinning the remaining steps so that
we have a set of effectively independent samples drawn from the posterior.
Quantitatively, this is determined by keeping only every $L$th sample after
burn-in, where $L$ is set as the length such that the normalized correlation
\begin{equation}
\frac{\text{Cov}[\theta_n^{(i)},\theta_n^{(i+L)}]}
{\text{Var}[\theta_n]}\le 0.01
\end{equation}
where $n$ is the index over the $N_p$ parameters and $(i)$ is the index
over the samples drawn.

We find the 95\% credible
limits in two-dimensional slices of our parameter space using
a total of 460,000 MCMC steps.  These are the solid
black contours in Figures \ref{fig:contours_gauss} through 
\ref{fig:contours_br} 
below.  They are referred to below as the ``final credible limits''
returned by MCMC.  To gauge
convergence as a function of time, we also plot the 95\% credible limit contours
taken from our set of four MCMC chains after they have sampled
a total of 100,000 points.  
These are the dotted green
contours in Figures \ref{fig:contours_gauss} through \ref{fig:contours_br}.
They are referred to as the ``pre-convergence'' credible limits.
Note that, even though MCMC has discovered its final value of $\chi^2_\text{min}$
by this point (see Figure \ref{fig:chimin}), this rapid convergence in
$\chi^2_\text{min}$ has not translated into convergence for the credible limits
on parameter space.

In the case of APS, the confidence limit is interpreted 
as that set of points found
which satisfy $\chi^2\le\chi^2_\text{lim}$.  Figures \ref{fig:contours_gauss}
and \ref{fig:contours_nn} plot these points after APS has sampled 100,000 
points in the case where $\chi^2_\text{lim}\equiv\chi^2_\text{min}+12.59$
using equations (\ref{eqn:covraw}) and (\ref{eqn:cov_nn}) as covariograms,
respectively.
We find that the confidence limits returned by APS overlap with
the final credible limits of MCMC, but are generally too permissive,
they include regions of parameter space which MCMC ultimately excludes.
It is tempting to say that
this is due to the fact that the $\chi^2_\text{min}$ discovered by
APS is $\sim3$ greater than the $\chi^2_\text{min}$ discovered by
MCMC (therefore, $\chi^2_\text{lim}$ is $\sim 3$ too large).
However, Figure \ref{fig:contours_br} uses the same data as
Figure \ref{fig:contours_gauss}, but plots only those APS points
such that $\chi^2\le 1283.3$, i.e. $\chi^2\le\chi^2_\text{min, MCMC}+12.59$.
Still we find that the APS confidence limit includes regions of parameter
space excluded by the final MCMC contour.  Thus, the difference in
parameter space constraints between APS and MCMC amounts to difference
between the frequentist confidence limit (see Appendix \ref{sec:frequentism})
and the Bayesian credible limit.  The fact that the two contraints overlap
significantly but not entirely indicates the value in performing multiple
statistical tests before drawing conclusions from one's data set.
It is also worth noting that in all cases, APS has thoroughly explored
the correct confidence limit in fewer samples than has MCMC.  Compare the
coverage of the red scatter plot to the green pre-convergence MCMC contours in
Figures \ref{fig:contours_gauss} through \ref{fig:contours_br}.

@@

\begin{figure*}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_100k_0_1.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_100k_0_2.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_100k_0_3.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_100k_0_4.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_100k_1_2.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_100k_1_3.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_100k_1_4.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_100k_2_3.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_100k_2_4.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_100k_3_4.eps}
}
\caption{
95\% confidence limits in two-dimensional slices of parameter space as
determined by MCMC and APS after 100,000 points have been sampled.  Solid (black)
contours are the baseline result found by 4 MCMC chains
after sampling a total of 460,000 points.  
Dotted (green) contours are the result found by a 4 MCMC
chains after each has sampled 25,000 points (for a total of 100,000
sampled points).  Red points show the $\chi^2\le\chi^2_\text{min}+12.59$
points found by APS after sampling a total of 100,000 points.
These plots use equation (\ref{eqn:covraw}) for the covariogram.
}
\label{fig:contours_gauss}
\end{figure*}

\begin{figure*}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_nn_100k_0_1.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_nn_100k_0_2.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_nn_100k_0_3.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_nn_100k_0_4.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_nn_100k_1_2.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_nn_100k_1_3.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_nn_100k_1_4.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_nn_100k_2_3.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_nn_100k_2_4.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_nn_100k_3_4.eps}
}
\caption{
95\% confidence limits in two-dimensional slices of parameter space as
determined by MCMC and APS after 100,000 points have been sampled.  Solid (black)
contours are the baseline result found by 4 MCMC chains
after sampling a total of 460,000 points.  
Dotted (green) contours are the result found by a 4 MCMC
chains after each has sampled 25,000 points (for a total of 100,000
sampled points).  Red points show the $\chi^2\le\chi^2_\text{min}+12.59$
points found by APS after sampling a total of 100,000 points.
These plots use equation (\ref{eqn:cov_nn}) for the covariogram.
}
\label{fig:contours_nn}
\end{figure*}
\begin{figure*}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_br_100k_0_1.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_br_100k_0_2.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_br_100k_0_3.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_br_100k_0_4.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_br_100k_1_2.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_br_100k_1_3.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_br_100k_1_4.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_br_100k_2_3.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_br_100k_2_4.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{scatter_br_100k_3_4.eps}
}
\caption{
The black and green contours are as in Figure \ref{fig:contours_gauss}.
The red points represent the $\chi^2\le 1283.3$ points sampled by the
APS run used to generate Figure \ref{fig:contours_gauss}.
}
\label{fig:contours_br}
\end{figure*}

>>>>

Gaussian processes use the sampled data $\{\vec{\theta}^{(i)},\chi^{2,(i)}\}$, 
where $i$ indexes over all of the sampled points, to predict
$\mu=\chi^{2,(q)}$ at the candidate point 
$\vec{\theta}^{(q)}$ by assuming that the function
$\vec{\theta}\rightarrow\chi^2$ represents a sample drawn from a random process 
distributed
across parameter space.  At each point in parameter space, 
$\chi^2$ is assumed to be
distributed according to a normal distribution with mean 
$\bar{\chi}^2$
and variance dictated by the ``covariogram''
$C_{ij}(\vec{\theta}^{(i)},\vec{\theta}^{(j)})$.
If we have $N_g$ measurements of $\chi^2$, then we assume
they are distributed according to
\begin{equation}
\label{eqn:likelihood}
P(\vec{\chi}^2)=
\frac{\exp\bigg[-\frac{1}{2}K^{-1}\sum_{i,j}^{N_g}
(\chi^{2,(i)}-\bar{\chi}^2)C^{-1}_{ij}
(\chi^{2,(j)}-\bar{\chi}^2)\bigg]}{\sqrt{2\pi}^{N_g}\text{det}|KC|^{N_g/2}}
\end{equation}
$C_{ij}$ is a function assumed by the user encoding how variations in
$\chi^2$ at one point in parameter space affect variations in 
$\chi^2$ at other
points in parameter space.  $K$ is a parameter controlling the
normalization of $C_{ij}$.
The diagonal elements $C_{ii}=1+\sigma^2_{ii}$ where 
$\sigma^2_{ii}$ is the intrinsic variance in the value
of $\chi^2$ at a single point $\vec{\theta}^{(i)}$.  
Rasmussen and Williams (2006) treat 
the special case
$\bar{\chi}^2=0$ and find (see their equations 2.19, 2.25 and 2.26)
\begin{eqnarray}
\mu&=&\sum_{i,j} C_{qi}C^{-1}_{ij}\chi^{2,(j)}\nonumber\\
\sigma^2&=&C_{qq}-\sum_{i,j}C_{qi}C^{-1}_{ij}C_{jq}\label{sig}
\end{eqnarray}
where the sums over $i$ and $j$ are sums over the sampled points 
in parameter space.  $C_{iq}$ relates the sampled point $i$ to the candidate
point $q$.
We do not wish to assume that the mean value of $\chi^2$ is zero everywhere.
Therefore, we modify the equation for $\mu$ to give
\begin{equation}
\label{mu}
\mu=\bar{\chi}^2+\sum_{i,j} C_{qi}C^{-1}_{ij}
(\chi^{2,(j)}-\bar{\chi}^2)
\end{equation}
where $\bar{\chi}^2$ is the algebraic mean of the 
sampled $\chi^{2,(i)}$.
Note the similarity to a multi-dimensional Taylor series expansion with the
covariance matrix playing the role of the derivatives.
Equation (\ref{sig}) differs from equation (6) in \cite{bryan} because
they used the semivariance
$$\gamma_{ij}=\text{var}[\chi^2(\vec{\theta}^{(i)})
-\chi^2(\vec{\theta}^{(j)})]$$
in place of the covariance $C_{ij}$.  In practice, the two assumptions result in
equivalently valid $\mu$ and $\sigma$.

@@

As mentioned above, because our toy $\chi^2$ function is constructed without any
parameter degeneracies, we are able to directly integrate the posterior and
determine the true Bayesian credible limits in our parameter space.  Figures
\ref{fig:toyBayes2}, \ref{fig:toyBayes3}, and \ref{fig:toyBayes4} plot the true
95\% credible limits in all 10 2-dimensional sub-spaces of our 5-dimensional
parameter space (red contours).  They also show the APS-determined credible
limits as described in Section \ref{sec:bayes} (black points).  The
APS-determined credible limits correspond well with the true credible limits,
giving us confidence in APS's ability to map out Bayesian credible limits as well
as Frequentist confidence limits.

The tests above demonstrate APS's ability to correctly determin credible and
confidence limits.  They do not, however, address the question of APS's chief
advantage over Markov Chain Monte Carlo methods: its ability to successfully
locate all of the low-$\chi^2$ regions in a given parameter space.  To
demonstrate this, we ran APS 100 times on each of the likelihood functions
plotted in Figures \ref{fig:toyFreq2}-\ref{fig:toyFreq4}.  Each time, the
pseudo-random number generator driving APS was given a different seed.  In all
cases, APS successfully located all of the local minima in $\chi^2$, usually in
fewer than 10,000 evaluations of $\chi^2$ (one case of the 2-mode $\chi^2$
function required 10,352 evaluations to find both low-$\chi^2$ regions; one
required 12,549 evaluations).  For comparison, we also ran 100 Markov Chain
Monte Carlo searches on these $\chi^2$ function, each run consisting of five
independent chains.  These searches were not nearly as successful as APS at
locating all of the low-$\chi^2$ regions.  In the case of the 2-mode function,
50 of the 100 MCMC searches failed to identify both modes.  In the case of the
3-mode function, 56 of the 100 MCMC searches failed to identify all of the
modes.  In the case of the 4-mode function, 87 of the 100 MCM searches failed to
identify all of the modes.  This is not to say that APS is perfect at
identifying all of the modes of a likelihood function.  If we reduce the
characteristic widths $w_{j,i}$ in equation (\ref{eqn:toychi}), we enter a regime
in which neither APS nor MCMC is capable of reliably,
identifying all of the modes.  However, this test does indicate that APS is more
robust than MCMC against multi-modal likelihood functions, as was first
demonstrated by Bryan {\it et al}. (2007).

@@

The parameter space we explore is the six-dimensional space of the spatially
flat concordance $\Lambda$CDM cosmology:
\begin{itemize}
\item$\Omega_bh^2$ -- the density of baryons in the Universe
\\
\item$\Omega_\text{CDM}h^2$ -- the density of cold dark matter in the Universe
\\
\item$\tau$ -- the optical depth to the surface of last scattering.  $\tau$ cannot be
constrained without polarizationd data, which we do not consider here.  However, it functions
as a nuisance parameter to be marginalized over.  The fact that APS can handle it as well as
MCMC is an important test of APS' feasibility.
\\
\item$h$ -- the Hubble parameter in units of
$100~\text{km}~\text{s}^{-1}~\text{Mpc}^{-1}$
\\
\item$n_s$ -- the spectral index controlling the distribution of scalar density
perturbations in the Universe
\\
\item$\ln[10^{10}A_s]$ -- the amplitude of the initial scalar density
perturbations in the Universe
\end{itemize}
These are the parameters taken by the publicly available MCMC code CosmoMC,
which we use to run our baseline MCMC chains \cite{cosmomc}.  The
$\vec{\theta}\rightarrow\chi^2$ function in this case involves converting these
paramters into the power spectrum of anisotropy $C_\ell$s on the sky and
comparing those $C_\ell$s to the actual measurements registered by the WMAP
satellite.  Using the (very fast) Boltzmann code CAMB to calculate the $C_\ell$s
\cite{camb} takes 1.3 seconds on a 2.5 GHz processor.  
Evaluating the $\chi^2$ using the likelihood code provided by the WMAP team
\cite{wmap7likelihood} takes an additional 2 seconds.  Hence the need
for an algorithm like APS or MCMC to accelerate the search.
We will consider both equations (\ref{eqn:GaussianCovar}) and (\ref{eqn:NNCovar})
as covariogram for our Gaussian Process, 
though we find that the choice has little effect on our
results.

@@

The second question we consider is whether or not our assumption that
$\vec{\theta}\rightarrow\chi^2$ can be modeled by a Gaussian Process is 
valid.  In Figure
\ref{fig:goodness} we plot the average fractional error in $\mu$ as a proxy for $\chi^2$ as a
function of the number of steps sampled (note: the vertical axis actually shows the average
fractional error over the preceding 500 steps so that the early, imprecise models do not
contaminate the later, precise models).  As APS samples more points, its Gaussian Process
model becomes more accurate.  This effect is more pronounced for covariogram (\ref{eqn:GaussianCovar})
than for covariogram (\ref{eqn:NNCovar}).  However, we will see below that this difference
does not effect the parameter constraints determined by APS.

@@

\begin{figure}
\includegraphics[scale=0.3]{goodness_of_fit.eps}
\caption{
The average value of $|\mu-\chi^2|/\chi^2$ as a function of the number of points
sampled by APS in the WMAP 7 problem.  
This is shown as a check on our assumption that
$\vec{\theta}\rightarrow\chi^2$ can be approximated as a Gaussian Process.
Though the squared exponential covariogram gives a better fit to the $\chi^2$ function,
we see in Figures \ref{fig:contours_gauss} and \ref{fig:contours_nn}
that this has little effect on the parameter constraints derived by APS.
}
\label{fig:goodness}
\end{figure}

We now consider the ultimate results, namely the 95\% confidence limits on
our 6-dimensional cosmological parameter space yielded by APS.
To provide a baseline against which to test APS, we run 4 independent MCMC
chains using CosmoMC.  Like the MCMC code used in Section \ref{sec:toy},
CosmoMC periodically adjusts its proposal density by learning the covariance
matrix of the points it has already sampled.  At each step, it proposes
a new point in parameter space by randomly selecting an eigen vector
of that learned covariance matrix and stepping along it.
We integrate the posterior by taking the output chains, discarding the first
50\% of steps as a burn in period, and thinning the remaining steps so that
we have a set of effectively independent samples drawn from the posterior.
Quantitatively, this is determined by keeping only every $L$th sample after
burn-in, where $L$ is set as the length such that the normalized correlation
\begin{equation}
\frac{\text{Cov}[\theta_n^{(i)},\theta_n^{(i+L)}]}
{\text{Var}[\theta_n]}\le 0.1
\end{equation}
where $n$ is the index over the $N_p$ parameters and $(i)$ is the index
over the samples drawn.

We find the 95\% credible
limits in two-dimensional slices of our parameter space using
a total of 460,000 MCMC steps.  These are assumed to be the
true credible limits on the parameter space and are plotted
as the thick red
contours in Figures \ref{fig:contours_gauss} through 
\ref{fig:contours_br} 
below.  Against these control contours, we plot the Frequentist
confidence limits (all points with $\chi^2\le\chi^2_\text{lim}$)
discovered by APS after only 50,000 evaluations of $\chi^2$.
We consider both an adaptive $\chi^2_\text{lim}=\chi^2_\text{min}+12.6$
and an absoulte $\chi^2_\text{lim}=1283.3$ (this is the 95\% confidence limit
for a $\chi^2$ probability distribution with 1199 degrees of freedom).
These are the black points in Figures \ref{fig:contours_gauss}, \ref{fig:contours_nn},
and \ref{fig:contours_br} respectively.  These confidence limits overlap very nicely
with the credible limits discovered by MCMC after many more $\chi^2$ evaluations,
comporting the the idea that, in the large data limit, Bayesiand credible limits
and Frequentist confidence limits ought to align.

For comparison of convergence properties, we plot the Bayesian credible limits discovered
by MCMC after only 50,000 $\chi^2$ evaluations.  These are the black regions in
Figure \ref{fig:mcmc_comparison}.  As can be seen, MCMC has not yet totally converged
in all of the sub-spaces of cosmological parameter space.  This should be compared
with the reasonably complete coverage of the control credible limits seen in
Figures \ref{fig:contours_gauss}-\ref{fig:contours_br}.

In Figure \ref{fig:contours_bayes}, we plot the 95\% Bayesian credible limit discovered
by APS after 50,000 evalutations of $\chi^2$ and determined by the method outlined in
Section \ref{sec:bayes}.  While we do not see the clean equivalence between the APS-derived credible limits and
the true credible limits that we saw in Figures \ref{fig:toyBayes2}-\ref{fig:toyBayes4},
we do find that the APS-derived credible limits roughly sketch the bounds of the 
true credible limit.  This, combined with its robustness against multi-modal likelihood
functions, suggests that APS may, at the very least be useful in cases where very little
prior information is available and exploring the paramter space is at least as important as
integrating it to find a rigorous bound.

@@

\begin{figure*}
\subfigure[]{
\includegraphics[scale=0.15]{gaussLearn_0_1_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussLearn_0_2_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussLearn_0_4_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussLearn_0_5_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussLearn_1_2_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussLearn_1_4_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussLearn_1_5_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussLearn_2_4_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussLearn_2_5_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussLearn_4_5_50k_frequentist.eps}
}
\caption{
The 10 2-dimensional sub-spaces of our 6-dimensional parameter space
($\tau$ is ignored because not real constraint can be gleaned from this data).
The thick red contours are the 95\% Bayesian credible limits determined by MCMC after
460,000 calls to $\chi^2$.  The black points are the 95\% Frequentist confidence limits
determined by APS after 50,000 calls to $\chi^2$.  $\chi^2_\text{lim}=\chi^2_\text{min}+12.6$.
Equation (\ref{eqn:GaussianCovar}) is used for the Gaussian Process covariogram.
}
\label{fig:contours_gauss}
\end{figure*}

\begin{figure*}
\subfigure[]{
\includegraphics[scale=0.15]{nnLearn_0_1_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{nnLearn_0_2_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{nnLearn_0_4_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{nnLearn_0_5_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{nnLearn_1_2_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{nnLearn_1_4_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{nnLearn_1_5_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{nnLearn_2_4_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{nnLearn_2_5_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{nnLearn_4_5_50k_frequentist.eps}
}
\caption{
The 10 2-dimensional sub-spaces of our 6-dimensional parameter space
($\tau$ is ignored because not real constraint can be gleaned from this data).
The thick red contours are the 95\% Bayesian credible limits determined by MCMC after
460,000 calls to $\chi^2$.  The black points are the 95\% Frequentist confidence limits
determined by APS after 50,000 calls to $\chi^2$.  $\chi^2_\text{lim}=\chi^2_\text{min}+12.6$.
Equation (\ref{eqn:NNCovar}) is used for the Gaussian Process covariogram.
}
\label{fig:contours_nn}
\end{figure*}

\begin{figure*}
\subfigure[]{
\includegraphics[scale=0.15]{gaussAssert_0_1_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussAssert_0_2_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussAssert_0_4_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussAssert_0_5_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussAssert_1_2_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussAssert_1_4_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussAssert_1_5_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussAssert_2_4_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussAssert_2_5_50k_frequentist.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussAssert_4_5_50k_frequentist.eps}
}
\caption{
The 10 2-dimensional sub-spaces of our 6-dimensional parameter space
($\tau$ is ignored because not real constraint can be gleaned from this data).
The thick red contours are the 95\% Bayesian credible limits determined by MCMC after
460,000 calls to $\chi^2$.  The black points are the 95\% Frequentist confidence limits
determined by APS after 50,000 calls to $\chi^2$.  $\chi^2_\text{lim}=1283.3$.
Equation (\ref{eqn:GaussianCovar}) is used for the Gaussian Process covariogram.
}
\label{fig:contours_br}
\end{figure*}

\begin{figure*}
\subfigure[]{
\includegraphics[scale=0.15]{mcmc_comparison_0_1_50k.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{mcmc_comparison_0_2_50k.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{mcmc_comparison_0_4_50k.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{mcmc_comparison_0_5_50k.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{mcmc_comparison_1_2_50k.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{mcmc_comparison_1_4_50k.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{mcmc_comparison_1_5_50k.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{mcmc_comparison_2_4_50k.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{mcmc_comparison_2_5_50k.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{mcmc_comparison_4_5_50k.eps}
}
\caption{
The 10 2-dimensional sub-spaces of our 6-dimensional parameter space
($\tau$ is ignored because not real constraint can be gleaned from this data).
The thick red contours are the 95\% Bayesian credible limits determined by MCMC after
460,000 calls to $\chi^2$.  The black regions are the 95\% Bayesian credible limits
determined by MCMC after 50,000 calls to $\chi^2$.
}
\label{fig:mcmc_comparison}
\end{figure*}

\begin{figure*}
\subfigure[]{
\includegraphics[scale=0.15]{gaussLearn_0_1_50k_bayes.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussLearn_0_2_50k_bayes.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussLearn_0_4_50k_bayes.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussLearn_0_5_50k_bayes.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussLearn_1_2_50k_bayes.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussLearn_1_4_50k_bayes.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussLearn_1_5_50k_bayes.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussLearn_2_4_50k_bayes.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussLearn_2_5_50k_bayes.eps}
}
\subfigure[]{
\includegraphics[scale=0.15]{gaussLearn_4_5_50k_bayes.eps}
}
\caption{
The 10 2-dimensional sub-spaces of our 6-dimensional parameter space
($\tau$ is ignored because not real constraint can be gleaned from this data).
The thick red contours are the 95\% Bayesian credible limits determined by MCMC after
460,000 calls to $\chi^2$.  The black points are the 95\% Bayesian credible limits determined
by APS according to the method outlined in Section \ref{sec:bayes}.  While the APS credible limit
covers the entire MCMC limit, it also covers regions of parameter space excluded by MCMC.
}
\label{fig:contours_bayes}
\end{figure*}
